---
title: "Getting started"
author: "Lilith Whittles"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
par(mar = c(3, 3, 1, 1), mgp = c(1.7, 0.7, 0), bty = "n")
```


## Vignette Info
First we will run the model without vaccination, read in your table of parameters
```{r}
 library("gonovax")
 filename <- system.file("extdata/gono_params_t.csv", package = "gonovax",
                         mustWork = TRUE)
 parameter_table <- read.csv(filename)
 parameter_table
```

We need to transform the table of parameters to the format the model requires
using the `transform_fixed` function. First we only use 10 parameter sets
```{r}

n_par <- 100
gono_params <- lapply(seq_len(n_par),
                      function(i) transform_fixed(parameter_table[i, ]))

# define the times of the output (in years)
tt <- c(0, 50)
y0 <- run_onevax_xvwv(tt, gono_params)

## gets final (equilibrium) states for restart
init_params <- lapply(y0, restart_params)

# generate vaccine effects parameters
ve <- data.frame(vea = rep(0.1, n_par),
                 ved = rep(0.2, n_par),
                 ves = rep(0.3, n_par))

tt <- seq(0, 10)
y <- run_onevax_xvwv(tt,
                     gono_params = gono_params,
                     init_params = init_params, 
                     vea = ve$vea, ved = ve$ved, ves = ve$ves,
                     uptake = 1, strategy = "VoA")
# strategies can be `VoA` vaccination on attendance,
# `VoD` = vaccination on diagnosis

```
`y` is now a list of model runs of length `n_par`. To interact easily with these
outputs you can use the `aggregate` function. 

This will aggregate over group (L, H) and strata (X, V, W) to give the total
cumulative infections for each parameter set (rows) over time (columns) 

```{r}

# 
total_infected <- aggregate(y, what = "cum_incid")
col <- rgb(0.5, 0.5, 0.5, 0.3)
matplot(tt, t(total_infected), lty = 1, type = "l", col = col,
        xlab = "Time", ylab = "Cumulative infections")
```

This will aggregate over group (L, H) and strata (X, V, W) to give the yearly
infections for each parameter set (rows) over time (columns) 

```{r}

# 
annual_infected <- aggregate(y, what = "cum_incid", as_incid = TRUE)
matplot(tt[-1], t(annual_infected), lty = 1, type = "l", col = col,
        xlab = "Time", ylab = "Annual infections")
```

You can look at individual vaccine strata:

```{r}

# 
annual_infected_X <- aggregate(y, what = "cum_incid", as_incid = TRUE,
                             stratum = "X")
annual_infected_V <- aggregate(y, what = "cum_incid", as_incid = TRUE,
                             stratum = "V")
col1 <- rgb(0.5, 0, 0, 0.3)
col2 <- rgb(0, 0, 0.5, 0.3)
matplot(tt[-1], t(annual_infected_X), lty = 1, type = "l", col = col1,
        xlab = "Time", ylab = "Annual infections")
matlines(tt[-1], t(annual_infected_V), lty = 1, col = col2)
legend("top", fill = c(col1, col2), legend = c("Unvaccinated", "Vaccinated"),
       ncol = 2)
```
You can look apply functions over the parameter sets to get summary statistics

```{r}


mean_ci <- function(x) c(mean = mean(x), quantile(x, c(0.025, 0.975)))
summary_annual_infected_X <- aggregate(y, what = "cum_incid", as_incid = TRUE,
                             stratum = "X", f = mean_ci)
summary_annual_infected_V <- aggregate(y, what = "cum_incid", as_incid = TRUE,
                             stratum = "V", f = mean_ci)
col1 <- "red"
col2 <- "blue"
matplot(tt[-1], t(summary_annual_infected_X), lty = c(1, 2, 2), type = "l", col = col1,
        xlab = "Time", ylab = "Annual infections")
matlines(tt[-1], t(summary_annual_infected_V), lty = c(1, 2, 2), col = col2)
legend("top", fill = c(col1, col2), legend = c("Unvaccinated", "Vaccinated"),
       ncol = 2)
```